---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r}
# Setup commands for this R-session:
# Code to set working directory (enabled):
if (TRUE) {
    working.directory <- paste0(Sys.getenv("HOME"), "/School/Coursera/DataScience_JohnsHopkins/05_Reproducable/Project/Project01")
    setwd(working.directory)
    #getwd()
}
```

## Loading and preprocessing the data

```{r}
filename <- "activity.csv"
all.data <- read.csv(filename, header=TRUE)

# Remove incomplete cases:
df.1 <- all.data[complete.cases(all.data),]

suppressPackageStartupMessages(library(xts))

# Create a tidy 'zoo' object:
datestamps <- as.POSIXct(df.1$date, format="%Y-%m-%d")
df.2 <- xts(cbind(df.1[, 3], df.1[, 1]), datestamps)
colnames(df.2) <- c("interval", "steps")
#class(df.2)
#colnames(df.2)
#head(df.2)
```

## What is mean total number of steps taken per day?

```{r}
steps.per.day <- apply.daily(df.2$steps, sum)

# Make a histogram of the total number of steps taken each day:
plot(steps.per.day, type="h", main="Steps Taken Each Day", xlab="Day", ylab="Steps")

# Calculate and report the mean and median total number of steps taken per day:
mean.steps.per.day <- sprintf("%s",floor(mean(steps.per.day)))
median.steps.per.day <- sprintf("%s",floor(median(steps.per.day)))
```

The mean total number of steps taken per day is `r mean.steps.per.day`.  
The median total number of steps taken per day is `r median.steps.per.day`.

## What is the average daily activity pattern?

```{r}
mean.steps.per.interval <- aggregate(steps ~ interval, FUN=mean, data=df.2)
median.steps.per.interval <- aggregate(steps ~ interval, FUN=median, data=df.2)

# Make a time series plot of the 5-minute interval (x-axis) and the average
# number of steps taken, averaged across all days (y-axis):
plot(mean.steps.per.interval, type="l", main="Average Number of Steps Taken per Day (1)", 
     xlab="Time (24-hour Time)", ylab="Steps")

# Which 5-minute interval, on average across all the days in the dataset,
# contains the maximum number of steps?:
max.mean.interval <- sprintf("%s", mean.steps.per.interval[which.max(mean.steps.per.interval$steps), 1])
max.mean.steps <- floor(mean.steps.per.interval[which.max(mean.steps.per.interval$steps), 2])
```

The `r max.mean.interval` 5-minute interval has the maximum average number of steps (`r max.mean.steps`).

## Imputing missing values

```{r}
# Calculate and report the total number of missing values in the dataset:
number.missing.values <- dim(all.data)[1] - dim(df.2)[1]

# Devise a strategy for filling in all of the missing values in the dataset:
# [JWT: Use the median value of the time interval.]
# Create a new dataset that is equal to the original dataset but with the missing
# data filled in:
df.3 <- all.data  # leave original data untouched
df.3$steps[is.na(df.3$steps)] <- median.steps.per.interval$steps[median.steps.per.interval$interval == df.3$interval[is.na(df.3$steps)]]

# Create a tidy 'zoo' object:
datestamps <- as.POSIXct(df.3$date, format="%Y-%m-%d")
df.3 <- xts(cbind(df.3[, 3], df.3[, 1]), datestamps)
colnames(df.3) <- c("interval", "steps")

mean.steps.per.interval <- aggregate(steps ~ interval, FUN=mean, data=df.3)
median.steps.per.interval <- aggregate(steps ~ interval, FUN=median, data=df.3)

# Make a histogram of the total number of steps taken each day:
plot(mean.steps.per.interval, type="l", main="Average Number of Steps Taken per Day (2)", 
     xlab="Time (24-hour Time)", ylab="Steps")

# Calculate and report the mean and median total number of steps taken per day:
mean.steps.per.day <- sprintf("%s",floor(mean(steps.per.day)))
median.steps.per.day <- sprintf("%s",floor(median(steps.per.day)))
```

The number of missing values in the data set is `r number.missing.values`.  
The mean total number of steps taken each day is `r mean.steps.per.day`.  
The median total number of steps taken each day is `r median.steps.per.day`.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
# Create a new factor variable in the dataset with two levels – “weekday”
# and “weekend”:
daytype <- ifelse(weekdays(index(df.3)) %in% c("Saturday", "Sunday"), "weekend", "weekday")
daytype <- as.factor(daytype)
df.4 <- merge(df.3, daytype)

mean.steps.per.interval <- aggregate(steps ~ interval + daytype, FUN=mean, data=df.4)

# Make a panel plot containing a time series plot of the 5-minute interval
# (x-axis) and the average number of steps taken, averaged across all weekday
# days or weekend days (y-axis):
suppressPackageStartupMessages(library(lattice))

xyplot(steps ~ interval | daytype, data=mean.steps.per.interval, type="l", layout=c(1,2),
       main="Average Number of Steps Taken per Day (3)",
       xlab="Time (24-hour Time)", ylab="Steps",
       strip=strip.custom(var.name="daytype", factor.levels=c("weekday","weekend"), strip.level=rep(TRUE,2)))
```

